---
title: "Fish diversity in mangroves"
author: "Jody Daniel"
date: "`r Sys.Date()`"
output:
 markdowntemplates::skeleton:
    highlight: tango
    toc: TRUE

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=TRUE, warning=FALSE, message=FALSE}
library(vegan)
library(here)
#install.packages(paste0(here::here(),"/packages/vegetarian_1.2.tar.gz"), repos = NULL, type = "source")
library(vegetarian)
library(readxl)
library(tidyverse)
library(indicspecies)
library(ggthemes)
library(skimr)
library(indicspecies)
source(paste0(here::here(), "/scripts/functions.R"))

```

```{r include=FALSE}
# import the data
fish_abundance_raw <- read_xlsx(here("data/Fish_MH-WH.xlsx"), sheet = 1)
fish_biomass_raw <- read_xlsx(here("data/Fish_MH-WH.xlsx"), sheet = 2)



```

The data are not following the a bell-shaped curve, and they are likely non-normal. Because the tests we will use make no assumption on the shape of the data, this is not really of concern

## Species Diversity

We first want to determine if there is a difference in diversity between the two sites. So, we will measure richness, shannon weiner and simpsons at the 2 sites and then run a t-test to determine if there is a difference.

```{r function for alpha, include=FALSE}

diversity_alpha <- function(x){
  library(vegetarian)
  A <- x[,3:ncol(x)]
  SiteInfo <- x[,1:2]
  Richness <- apply(A, 1, function(x){H(x, lev = "alpha", q = 0)})
  Shannon <- apply(A, 1, function(x){H(x, lev = "alpha", q = 1)})
  Simpson <- apply(A, 1, function(x){H(x, lev = "alpha", q = 2)})
  Diversity <- tibble(SiteInfo,Richness,`Shannon-Weiner` = Shannon,Simpson)
  return(Diversity)
}


fish_alpha <- diversity_alpha(fish_abundance_raw)

```

```{r test for alpha}
# is there a difference between the two sites?

t.test(Richness ~ Site, data = fish_alpha) # is there a difference in richness? - yes
t.test(`Shannon-Weiner` ~ Site, data = fish_alpha) # is there a difference in shannon weiner - yes
t.test(Simpson ~ Site, data = fish_alpha) # is there a difference in simpsons - yes? if we use an alpha of 0.1

```

## Is there a difference in comuunity composiiton?

Now, we will use a mrpp to see if there is a difference in community composition between the two sites.I used an mrpp because after removing sites with no observations, we ended up with an unbalanced design. 

```{r test for community comp}
rows_empty <- which(rowSums(fish_abundance_raw[, 3:ncol(fish_abundance_raw)])>0)# some replicates have no observations
mrpp(fish_abundance_raw[rows_empty,3:ncol(fish_abundance_raw)], fish_abundance_raw$Site[rows_empty]) # is there a difference in community composition based on abundance? - no: p-value = 0.127, A (which is the test statistic) = 0.0007263 (values closer to 1 means there is a large difference, values closer to zero means there is little difference) 
mrpp(fish_biomass_raw[rows_empty,3:ncol(fish_biomass_raw)], fish_biomass_raw$Site[rows_empty])# is there a difference in community composition based on biomass ? - no : p-value = 0.458, A (which is the test statistic) = -0.0004895  (values closer to 1 means there is a large difference, values closer to zero means there is little difference) 
```

## Are some species more associated with one site over the other the other?

We will use an indicator species analysis.

```{r isa, echo=FALSE}

fish_abundance_isa <- multipatt(x = fish_abundance_raw[, 3:ncol(fish_abundance_raw)], 
                                cluster = fish_abundance_raw$Site)
fish_abundance_isa$sign # no species have a significant p_value, so none are more strongly associated with one site over the other

fish_biomass_isa <- multipatt(x = fish_biomass_raw[, 3:ncol(fish_biomass_raw)], 
                                 cluster = fish_biomass_raw$Site)
fish_biomass_isa$sign # no species have a significant p_value, so none are more strongly associated with one site over the other
fish_abundance_isa$B

```
## Making figures


```{r include=FALSE}
#theme_tufte
fish_alpha_long <- 
  fish_alpha %>%
  gather(key = "Alpha", value = "Value", Richness, `Shannon-Weiner`, Simpson)%>%
  mutate(Significance = case_when(Alpha == "Richness" ~ "***",
                                 Alpha == "Shannon-Weiner" ~ "**",
                                 Alpha == "Simpson" ~ "*"))
FishSig <- fish_alpha_long %>% group_by(Site, Alpha, Significance) %>% 
  summarise(Value = mean(Value)) 

 

png(here("outputs/FishDiversity.png"), width = 8, height = 5,
    res = 600, units = "in")
ggplot()+
  stat_summary(data = fish_alpha_long, aes(x = Site, y = Value, 
                                           fill = Site),
               fun = "mean", geom = "bar", position = "dodge")+
  stat_summary(data = fish_alpha_long, aes(x = Site, y = Value),
               fun.data = "mean_se", geom = "errorbar", position = "dodge")+
  geom_text(data = FishSig,
            aes(x = Site, y = Value+0.25,
                label = Significance), size = 7.5, position = position_dodge(width = .9))+
  labs(y = "Alpha Diversity", x = " ",
       title = " ", fill = " ")+
  scale_x_discrete(labels = c("Mount Hartman",
                              "Westerhall"))+
 scale_fill_viridis_d(option = "magma",
                      labels = c("Mount Hartman",
                              "Westerhall"))+
      scale_color_viridis_d(option = "magma")+
      theme_bw()+
      theme(
        # X-axis title: no angle, all the way to the right
        axis.title.x = element_text(angle = 0, hjust = 1, vjust = 0),
        
        # Y-axis title: normal vertical orientation, all the way to the left
        axis.title.y = element_text(angle = 90, hjust = 1, vjust = 0),
        
        legend.position = "none")+
  facet_grid(~Alpha)
dev.off()
```


```{r include=FALSE}
#theme_tufte
fish_alpha_long <- 
  fish_alpha %>%
  gather(key = "Alpha", value = "Value", Richness, `Shannon-Weiner`, Simpson)%>%
  mutate(Significance = case_when(Alpha == "Richness" ~ "***",
                                 Alpha == "Shannon-Weiner" ~ "**",
                                 Alpha == "Simpson" ~ "*"))


 

png(here("outputs/FishDiversityJCO.png"), width = 8, height = 5,
    res = 600, units = "in")
ggplot()+
  stat_summary(data = fish_alpha_long, aes(x = Alpha, y = Value),
               fun = "mean", geom = "bar", position = "dodge")+
  stat_summary(data = fish_alpha_long, aes(x = Alpha, y = Value),
               fun.data = "mean_se", geom = "errorbar", position = "dodge")+
  labs(x = "Alpha Diversity", y = "Value",
       title = " ", fill = " ")+
      scale_fill_viridis_d(option = "magma")+
      theme_bw()+
      theme(
        # X-axis title: no angle, all the way to the right
        axis.title.x = element_text(angle = 0, hjust = 1, vjust = 0),
        
        # Y-axis title: normal vertical orientation, all the way to the left
        axis.title.y = element_text(angle = 90, hjust = 1, vjust = 0),
        
        legend.position = "none")
dev.off()
```



```{r}
png(here("outputs/FishGuildJCO.png"), width = 8, height = 5,
    res = 600, units = "in")
FishDictionary<-
  read_excel(paste0(here::here(), "/data/DataForFigShare.xlsx"), sheet = 2)

fish_abundance_raw %>% select(-Site, -Replicate)%>%
  pivot_longer(ATF:FF, names_to = "Species", values_to = "Abundance")%>%
  left_join(FishDictionary, by = c("Species" = "SpeciesCode"))%>%
  filter(!is.na(SpeciesName))%>%
   group_by(`FunctionalGuild/Trait`)%>%
  reframe(Abundance = sum(Abundance))%>%
ggplot(aes(x = `FunctionalGuild/Trait`, y = Abundance))+
  geom_bar(stat = "identity")+
      scale_fill_viridis_d(option = "magma")+
      theme_bw()+
  labs(x = "Functional Guild/Trait")+
      theme(
        # X-axis title: no angle, all the way to the right
        axis.title.x = element_text(angle = 0, hjust = 1, vjust = 0),
        
        # Y-axis title: normal vertical orientation, all the way to the left
        axis.title.y = element_text(angle = 90, hjust = 1, vjust = 0),
        
        legend.position = "none")
dev.off()
```
